---
title: "Airports"
author: "Namita"
date: "August 9, 2019"
output: html_document
---

# Flights at ABIA

Dataset contains all incoming and outgoing Flights flying in and out of the Austin-Bergstrom Interational Airport. 

## Reading in the data
```{r echo = FALSE, include = FALSE}
library(ggplot2)
library(ggthemes)
library(directlabels)
library(sqldf)
library(tidyverse)
library(ggmap)
library(dplyr)
library(maps)
library(geosphere)
library(RColorBrewer)
#rm(list=ls())
setwd("C:/Users/namit/Desktop/MSBA/Intro to Predictive Analytics/STA380 - The journey continues/")

# Read ABIA data and airport codes
abia <- read.csv("ABIA.csv",header=TRUE, sep=",")
airport_codes <- read.csv("airport-codes.csv")

#Splitting into 2 separate dataframes based on destination/source
attach(abia)
colnames(abia)[colnames(abia)=="ï..Year"] <- "Year"
dfFromAUS <- abia[which(abia$Origin == "AUS"),]
dfToAUS <- abia[which(abia$Dest == "AUS"),]

```
Here's a peak into the dataset:
1. 99260 rows of data mapped to 29 columns  
2. Missing values - It can be observed from the table given below that the number of missing values for certain columns such as CarrierDelay,WeatherDelay,NASDelay,SecurityDelay,LateAircraftDelay are way too high (i.e 79513). Hence it may not be very helpful to use these in our analysis.  

```{r echo = FALSE, warning = FALSE}
dim(abia)
missing_values <- as.data.frame(sapply(abia, FUN = function(x){sum(is.na(x))}))
missing_values$Columns <- rownames(missing_values)
rownames(missing_values) <- NULL
names(missing_values) <- c("Missing_Count", "Column Name")
missing_values <- missing_values[c(2,1)]
missing_values
```
# Overview of Insights
Here's a list of interesting insights we'd like to gather from the dataset:  
1. What are the top 5 destinations that people tend to fly to from Austin?   
2. What times of the year get most traffic to the top 5 destinations from Austin?  
3. Arrival and Departure Delays by Carrier  
4. Delays by time of the day vs Month  
5. Visualization of incoming and outgoing Air traffic from ABIA  


### Top 5 Destinations from Austin  
The top 5 destinations people fly out to from Austin are Dallas, Dallas(again!), Houston, Phoenix and Denver. There are over 11000 flights from Austin to Dallas in 2008, which is almost a staggering 900 outgoing flights to Dallas from Austin!  
Dallas-Forth Worth International Airport (DFW) serves as headquarters to the world’s largest carrier, American Airlines, while Dallas Love (DAL) is home to the world’s biggest low-cost carrier, Southwest. Dallas could arguably be considered the most important city in the U.S. aviation industry.   
```{r echo = FALSE, warning = FALSE}
top_DEST_from_AUS <- sqldf("select Dest, count(*) as 'Num_of_Flights' from dfFromAUS where Cancelled == 0 group by Dest order by Num_of_Flights DESC ")
top_DEST_from_AUS %>% head(5) %>%
   ggplot(aes(x = reorder(Dest,Num_of_Flights), y = Num_of_Flights, fill = Num_of_Flights)) + 
  geom_bar(stat ='identity') + 
  coord_flip() + 
  scale_x_discrete("Destination Airports") +
  scale_y_continuous("Number of Flights in 2008") +
  ggtitle("Top 5 Destinations from Austin") +
  theme_minimal()

top_5_dest <- head(top_DEST_from_AUS,5)
top_5_dest <- as.character(top_5_dest$Dest)
top_5_dest

```
### Monthwise trends in departures to Top 5 Destinations  
Digging deeper to see if there are any interesting trends in frequency of flights to these top 5 destinations, we found that there was a noticeable change in the number of departures to the aiports in Dallas - DFW and DAL. This was peculiar because we would expect the distribution to be remotely similar to both these airports with the end destinations being the same!  
What is important to note here is the period in context. 2008 was a tough year for global economies. The impact on businesses was also evident. We figured that 2008, especially the latter part of the year, must have been a tough year for Southwest Airlines (DAL is home to this low-cost budget airlines company). Our intuition turned out to be true as can be seen in this interview with the CEO from 2008 - https://www.sfgate.com/business/article/Southwest-CEO-describes-2008-s-challenges-3292577.php   

```{r echo = FALSE, warning = FALSE}
month_data <- abia %>% filter(Origin == "AUS", Dest %in% top_5_dest, Cancelled == 0) %>% group_by(Dest, Month) %>% summarise("Num_of_Flights" = n()) %>% arrange(-Num_of_Flights)%>%ggplot(aes(x = as.factor(Month), y = Num_of_Flights, group = Dest,color=Dest)) + 
  geom_line(aes(linetype = Dest)) +
  theme_minimal() +
  geom_dl(aes(label = Dest), method = list("last.points", hjust = 0.5, vjust = -0.5)) +
  scale_y_continuous("Number of flights")+
  scale_x_discrete("Month") +
  ggtitle("Departures per month in 2008 to top 5 destinations")

month_data

```

### Delay Analysis  
"Research in consumer psychology shows that customers seek reasons for service failures and that attributions of blame moderate the effects of failure on the level of customer satisfaction" - Quoted in a journal titled "The Impact of Service Operations Failures on
Customer Satisfaction: Evidence on How Failures and
Their Source Affect What Matters to Customers" (https://pdfs.semanticscholar.org/1487/dc65b163be21554246aac9b9884c7916d671.pdf)


Delays in arrival or departure thus contribute significantly to customer experience while flying in a particular airline. Thus, we decided to look into average delays in departure and arrival.  
```{r echo = FALSE, warning = FALSE}
#Plotting Dep delays by Carriers
library(dplyr)
library(ggplot2)
library(sqldf)
df1 <- select(dfFromAUS, UniqueCarrier, DepDelay)
res = summarise_at(group_by(df1,UniqueCarrier),vars(DepDelay),funs(mean(.,na.rm=TRUE)))
ylim <- c(0, 1.1*max(res$DepDelay))
x1 <- barplot(res$DepDelay ~ res$UniqueCarrier, data = res, width = 1, main = "Dep Delays", xlab="Carriers", ylab = "DepDelay (in minutes)", las = 2, ylim=ylim) 
text(x = x1, y = round(res$DepDelay), label = round(res$DepDelay), pos = 3, cex = 0.8, col = "red")

#Delay Density
delayDensity <- sqldf("select UniqueCarrier, sum(DepDelay)/count(*) as 'Delay_Density' from abia where DepDelay > 0 group by UniqueCarrier ;")

ylim <- c(0, 1.1*max(delayDensity$Delay_Density))
DepDensity <- barplot(delayDensity$Delay_Density ~ delayDensity$UniqueCarrier, data = res, width = 1, main = "Density of Dep Delays", xlab="Carriers", ylab = "DepDelay (in minutes)/No. of flights", las = 2, ylim=ylim) 

```


```{r echo = FALSE, warning = FALSE}
#Plotting Arrival delays by Carriers

df2 <- select(dfToAUS, UniqueCarrier, ArrDelay)
out = summarise_at(group_by(df2,UniqueCarrier),vars(ArrDelay),funs(mean(.,na.rm=TRUE)))
ylim <- c(0, 1.1*max(out$ArrDelay))
x2 <- barplot(out$ArrDelay ~ out$UniqueCarrier, data = out, width = 1, main = "Arr Delays", xlab="Carriers", ylab = "Arrival Delay (in minutes)", las = 2, ylim=ylim) 
text(x = x2, y = round(out$ArrDelay), label = round(out$ArrDelay), pos = 3, cex = 0.8, col = "red")

#Arrival Delay Density
ArrdelayDensity <- sqldf("select UniqueCarrier, sum(ArrDelay)/count(*) as 'Arr_Delay_Density' from abia where ArrDelay > 0 group by UniqueCarrier ;")

ylim <- c(0, 1.1*max(ArrdelayDensity$Arr_Delay_Density))
ArrDensity <- barplot(ArrdelayDensity$Arr_Delay_Density ~ ArrdelayDensity$UniqueCarrier, data = res, width = 1, main = "Density of Arrival Delays", xlab="Carriers", ylab = "Arrival Delay (in minutes)/No. of flights", las = 2, ylim=ylim) 
```

First we plotted the Delay in minutes for both Departure and Arrival with respect to the Carrier. However, the data was not conclusive. This was because the number of flights per airlines in this given dataset varies considerably. Hence to get a more accurate representation of the airline carriers who tend to have the most delays, we calculated the delay densities of each airline by computing the total delay by the number of flights departed/arrived for each carrier.  

### Delay by time of the day  

To find out the times of the day when delays are most likely, we bucketed the Scheduled Departure time into 7 buckets as follows:  
1. Early Morning (12am - 6am)  
2. Morning (6am - 9am)  
3. Pre-Afternoon (9am - 12pm)  
4. Affternoon (12pm - 3pm)  
5. Evening (3pm - 6pm)  
6. Night (6pm- 9pm)  
7. Late Night (9pm - 12am)  

We are now plotting the most likely zones of delay in a day by month.  
```{r echo = FALSE, warning = FALSE}
abia$CRSDepTimeCategory <- NA
abia$CRSDepTimeCategory[abia$CRSDepTime <= 600] <- "Early Morning"
abia$CRSDepTimeCategory[abia$CRSDepTime > 600 & abia$CRSDepTime <= 900] <- "Morning"
abia$CRSDepTimeCategory[abia$CRSDepTime > 900 & abia$CRSDepTime <= 1200] <- "Pre-Afternoon"
abia$CRSDepTimeCategory[abia$CRSDepTime > 1200 & abia$CRSDepTime <= 1500] <- "Afternoon"
abia$CRSDepTimeCategory[abia$CRSDepTime > 1500 & abia$CRSDepTime <= 1800] <- "Evening"
abia$CRSDepTimeCategory[abia$CRSDepTime > 1800 & abia$CRSDepTime <= 2100] <- "Night"
abia$CRSDepTimeCategory[abia$CRSDepTime > 2100 & abia$CRSDepTime <= 2359] <- "Late Night"
abia$CRSDepTimeCategory <- factor(abia$CRSDepTimeCategory,
                                        levels = c("Early Morning", "Morning", "Pre-Afternoon",
                                                   "Afternoon", "Evening", "Night", "Late Night"),
                                     ordered = TRUE)

flights_delayed <- as.data.frame(abia %>%
                                   filter(Cancelled==0, DepDelay > 30, Origin == "AUS") %>%
                                   group_by(Month, CRSDepTimeCategory) %>%
                                   summarise("Flights_Delayed" = n()))
total_flights <- as.data.frame(abia %>%
                                   filter(Cancelled==0, Origin == "AUS") %>%
                                   group_by(Month, CRSDepTimeCategory) %>%
                                   summarise("Total_Flights" = n()))
merge_df <- left_join(total_flights, flights_delayed)
merge_df$Prop_Delayed <- merge_df$Flights_Delayed*100/merge_df$Total_Flights

myColors <- brewer.pal(3, "Blues")

ggplot(data = merge_df, 
       aes(x = as.factor(Month), y = CRSDepTimeCategory, fill = Prop_Delayed)) + 
  geom_tile() +
  scale_fill_gradientn('Scale of Flights Delayed', colors = myColors) +
  theme_minimal() + 
  ggtitle("Delay times by Month") +
  scale_y_discrete("Departure Times") +
  scale_x_discrete("Month")
```
It can be seen from the graph above that the upper half of the graph depecting the latter parts of a day have more darker shaded squares implying that _**more delays happen in the second half of the day**_.
_Highest delays can be observed in the month of December_ given the obvious weather conditions, especially at Night.  


### Visualization depicting Flights out of Austin
```{r echo = FALSE, warning = FALSE}

#rm(list=ls())
setwd("C:/Users/namit/Desktop/MSBA/Intro to Predictive Analytics/STA380 - The journey continues/")
library(tidyverse)
library(ggmap)
library(dplyr)
library(maps)
library(geosphere)

## Read flight and airports lists 
abia <- read.csv("ABIA.csv",header=TRUE, sep=",")
airports_file <- "airports.csv"
if (file.exists(airports_file)) {
  airports <- read.csv(airports_file)
  } else {
  airports <- data_frame(airport = NA, lon = NA, lat= NA)
}

attach(abia)
# Rename a column in R
colnames(airports)[colnames(airports)=="ï..Airport"] <- "Airport"
attach(abia)
colnames(abia)[colnames(abia)=="ï..Year"] <- "Year"
dfFromAUS <- abia[which(abia$Origin == "AUS"),]

```

```{r echo = FALSE, warning = FALSE}
flights_from_aus <- merge(dfFromAUS, airports, by.x = "Dest", by.y = "Airport")
flights_from_aus <- merge(flights_from_aus, airports, by.x = "Origin", by.y = "Airport")
flights_from_aus <- flights_from_aus %>% 
  select(Origin, Dest, lon.x, lat.x, lon.y, lat.y) %>% 
  as_data_frame()

#create basemap
map("world", regions=c("usa"), fill=T, col="grey8", bg="grey15", ylim=c(21.0,50.0), xlim=c(-130.0,-65.0))
#overlay airports
points(airports$lon,airports$lat, pch=4, cex=0.4, col="chocolate1")

#Plot route lines
for (i in (1:dim(flights_from_aus)[1])) { 
inter <- gcIntermediate(c(flights_from_aus$lon.x[1], flights_from_aus$lat.x[1]), c(flights_from_aus$lon.x[i], flights_from_aus$lat.x[i]), n=200)
lines(inter, lwd=0.1, col="turquoise2")    
}

```