---
title: "Portfolio Modeling"
output: github_document
---

### Loading all required libraries  

```{r warning=FALSE}
library(mosaic)
library(quantmod)
library(foreach)
```

### 1. Importing top 3 oil & gas ETF's - last 5 yrs data - Portfolio 1  
### 2. Importing top 3 Building&Construction & top 2 Commodities ETF's - last 5 yrs data - Portfolio 2  
### 3. Import top 7 Diversified Portfolio ETF's - last 5 yrs data - Portfolio 3  

### Creating an all return matrix for all the 3 portfolios  

```{r warning=FALSE, error=FALSE}

# Import top 3 oil & gas ETF's - last 5 yrs data
etf1 = c("USO", "UNG", "DBO")

# Import top 3 Building&Construction & top 2 Commodities ETF's - last 5 yrs data
etf2 = c("ITB", "XHB", "PKB","PDBC","DBC")

# Import top 7 Diversified Portfolio ETF's - last 5 yrs data
etf3 = c("AOR", "AOM", "AOA", "MDIV", "AOK", "IYLD", "GAL")

#Getting data for past 5 years
getSymbols(etf1, from = "2014-08-01")
getSymbols(etf2, from = "2014-08-01")
getSymbols(etf3, from = "2014-08-01")


for(ticker in etf1) {
	expr = paste0(ticker, "a = adjustOHLC(", ticker, ")")
	eval(parse(text=expr))
}

all_returns1 = cbind(	ClCl(USOa),
								ClCl(UNGa),
								ClCl(DBOa))

# Creating all return matrix for Portfolio 1
all_returns1 = as.matrix(na.omit(all_returns1))
summary(all_returns1)

for(ticker in etf2) {
	expr = paste0(ticker, "a = adjustOHLC(", ticker, ")")
	eval(parse(text=expr))
}

all_returns2 = cbind(	ClCl(ITBa),
								ClCl(XHBa),
								ClCl(PKBa),
								ClCl(PDBCa),
								ClCl(DBCa))

# Creating all return matrix for Portfolio 2
all_returns2 = as.matrix(na.omit(all_returns2))
summary(all_returns2)

for(ticker in etf3) {
	expr = paste0(ticker, "a = adjustOHLC(", ticker, ")")
	eval(parse(text=expr))
}

all_returns3 = cbind(	ClCl(AORa),
								ClCl(AOMa),
								ClCl(AOAa),
								ClCl(MDIVa),
								ClCl(AOKa),
								ClCl(IYLDa),
								ClCl(GALa))
# Creating all return matrix for Portfolio 3
all_returns3 = as.matrix(na.omit(all_returns3))
summary(all_returns3)
```

### Calculating VAR for Portfolio 1  
#### From the summary of returns, it seems ETF "UNG" has a neagtive mean and median return. Hence assigning weights as (0.4,0.2,0.4) to have lesser weights on "UNG"  

```{r}

# Monte Carlo Simulation to Calculate VAR for Portfolio 1
initial_wealth = 100000
# 5000 simulations
sim1 = foreach(i=1:5000, .combine='rbind') %do% {
	total_wealth = initial_wealth
	weights = c(0.4, 0.2, 0.4)
	holdings = weights * total_wealth
	n_days = 20
	wealthtracker1 = rep(0, n_days)
	for(today in 1:n_days) {
		return.today = resample(all_returns1, 1, orig.ids=FALSE)
		holdings = holdings + holdings*return.today
		total_wealth = sum(holdings)
		wealthtracker1[today] = total_wealth
		holdings = weights * total_wealth #re-balancing each day
		
	}
	((wealthtracker1[20]-initial_wealth)*100)/initial_wealth
	
}
sorted_sim_1=sort(sim1)
VAR1=sorted_sim_1[250]
VAR1

```

### Calculating VAR for Portfolio 2  
#### From the summary of returns, it seems ETF "PDBC" and "DBC" hvae neagtive mean returns. Hence assigning the weights accordingly as (0.2,0.3,0.3,0.1,0.1) to have lesser weights on "PDBC" and "DBC"  

```{r}

# Monte Carlo Simulation to Calculate VAR for Portfolio 2
initial_wealth = 100000
#5000 Simulations
sim2 = foreach(i=1:5000, .combine='rbind') %do% {
	total_wealth = initial_wealth
	weights = c(0.2, 0.3, 0.3, 0.1, 0.1)
	holdings = weights * total_wealth
	n_days = 20 #20 trading day
	wealthtracker2 = rep(0, n_days)
	for(today in 1:n_days) {
		return.today = resample(all_returns2, 1, orig.ids=FALSE)
		holdings = holdings + holdings*return.today
		total_wealth = sum(holdings)
		wealthtracker2[today] = total_wealth
		holdings = weights * total_wealth #re-balancing each day
		
	}
	((wealthtracker2[20]-initial_wealth)*100)/initial_wealth
	
}
sorted_sim_2=sort(sim2)
VAR2=sorted_sim_2[250]
VAR2

```

### Calculating VAR for Portfolio 3  
#### From the summary of returns, all ETF's have comparable returns. Hence assigning almost equal weights to them.  

```{r}

# Monte Carlo Simulation to Calculate VAR for Portfolio 3
initial_wealth = 100000
#5000 simulations
sim3 = foreach(i=1:5000, .combine='rbind') %do% {
	total_wealth = initial_wealth
	weights = c(0.15, 0.15, 0.15, 0.15,0.15,0.15,0.1)
	holdings = weights * total_wealth
	n_days = 20
	wealthtracker3 = rep(0, n_days)
	for(today in 1:n_days) {
		return.today = resample(all_returns3, 1, orig.ids=FALSE)
		holdings = holdings + holdings*return.today
		total_wealth = sum(holdings)
		wealthtracker3[today] = total_wealth
		holdings = weights * total_wealth #re-balancing each day
		
	}
	((wealthtracker3[20]-initial_wealth)*100)/initial_wealth
	
}
sorted_sim_3=sort(sim3)
VAR3=sorted_sim_3[250]
VAR3

```

```{r}
print("Final Report")
cat("\nPortfolio 1 containing top 3 oil & gas ETF's - ",etf1, "\n4-week (20 trading day) VAR at the 5% level = ", VAR1,"%")
cat("\n\nPortfolio 2 containing top 3 Building&Construction & top 2 Commodities ETF's - ",etf2, "\n4-week (20 trading day) VAR at the 5% level = ", VAR2,"%")
cat("\n\nPortfolio 3 containing top 7 Diversified Portfolio ETF's - ",etf3, "\n4-week (20 trading day) VAR at the 5% level = ", VAR3,"%")
```

### **Final Report**


Portfolio  | ETF List                                          | ETF Category                                     | 4-week VAR
---------  | ---------------                                   | ------------------                               | -------------
 1         | "USO", "UNG", "DBO"                               | Top 3 oil & gas                                  | -14.64 %
 2         | "ITB", "XHB", "PKB","PDBC","DBC"                  | Top 3 Building&Construction & top 2 Commodities  |  -6.68 %
 3         | "AOR", "AOM", "AOA", "MDIV", "AOK", "IYLD", "GAL" | Top 7 Diversified Portfolios                     |  -2.79 %