---
title: "Visual story telling part 1: Green Buildings"
output: github_document
---

```{r echo=FALSE, results='hide',message=FALSE}
rm(list=ls())

library(mosaic)
library(tidyverse)



Building_Data = read.csv(file="/Users/nithi/Downloads/STA380-master/STA380-master/data/greenbuildings.csv", header=TRUE, sep=",")


```

### Let us take a look at the Data by running str(Building_Data). Also ran is.na(Building_Data) to take a look at any null values. We see that empl_gr has 71 null values.  


```{r echo=FALSE,message=FALSE}

str(Building_Data)
summary(is.na(Building_Data))


```



### We see that the leasing rate is 0 for many buildings which is weird since these are the buildings with less occupancy. Let us remove all these buildings with Leasing Rate less than 15%.  


```{r echo=FALSE, message=FALSE}
ggplot(data=Building_Data, aes(Building_Data$leasing_rate)) + 
  geom_histogram() +
geom_vline(data=Building_Data, aes(xintercept = median(Building_Data$leasing_rate)), colour="red") 


Building_Data_removed_15=Building_Data[which(!Building_Data$leasing_rate < 15), ] 


ggplot(data=Building_Data_removed_15, aes(Building_Data_removed_15$leasing_rate)) + 
  geom_histogram() +
geom_vline(data=Building_Data_removed_15, aes(xintercept = median(Building_Data_removed_15$leasing_rate)), colour="red") 

```


```{r }

median(Building_Data$leasing_rate)

median(Building_Data_removed_15$leasing_rate)

```
```{r echo=FALSE, message=FALSE}
Building_Data=Building_Data[which(!Building_Data$leasing_rate < 15), ] 
```


### Now let us consider green and non-green buildings seperately  
### We see that the size is not equal among the data, hence we cannot take this data to perform our analysis. 


```{r echo=FALSE, message=FALSE}

green_buildings=Building_Data[which(Building_Data$green_rating == 1), ] 

ggplot(data=green_buildings, aes(green_buildings$size)) + 
  geom_histogram() +
geom_vline(data=green_buildings, aes(xintercept = median(green_buildings$size)), colour="red") 



non_green_buildings=Building_Data[which(!Building_Data$green_rating == 1), ] 

ggplot(data=non_green_buildings, aes(non_green_buildings$size)) + 
  geom_histogram() +
geom_vline(data=non_green_buildings, aes(xintercept = median(non_green_buildings$size)), colour="red") 



print(c('The median size for green buildings is : ',median(green_buildings$size)))

print(c("The median size for non-green buildings is : ", median(non_green_buildings$size)))

```


### It is not accurate to calculate the median rent value from this data since it considers all buildings.  
### We should consider buildings only with size close to 250,000 and stories close to 15. Also we should consider the fact that the building will be constructed near downtown. Hence must consider only clusters in downtown.  
### With this in mind, let us shrink the data to obtain buildings in clusters with median height to be around 250,000 and stories 15.   


```{r echo=FALSE, message=FALSE}
ggplot(data=green_buildings, aes(green_buildings$stories)) + 
  geom_histogram() +
geom_vline(data=green_buildings, aes(xintercept = median(green_buildings$stories)), colour="red")


ggplot(data=non_green_buildings, aes(non_green_buildings$stories)) + 
  geom_histogram() +
geom_vline(data=non_green_buildings, aes(xintercept = median(non_green_buildings$stories)), colour="red")



print("The median stories for green buildings is : ")

median(green_buildings$stories)


print("The median stories for non-green buildings is : ")

median(non_green_buildings$stories)



```



```{r echo=FALSE, message=FALSE}



ggplot(data=green_buildings, aes(green_buildings$Rent)) + 
  geom_histogram() +
geom_vline(data=green_buildings, aes(xintercept = median(green_buildings$Rent)), colour="red") 

ggplot(data=non_green_buildings, aes(non_green_buildings$Rent)) + 
  geom_histogram() +
geom_vline(data=non_green_buildings, aes(xintercept = median(non_green_buildings$Rent)), colour="red") 



print("The median rent for green buildings is : ")

median(green_buildings$Rent)


print("The median rent for non-green buildings is : ")

median(non_green_buildings$Rent)

```

### We cannot consider this graph where the median Rent is 25 since it is for the entire data. Let us slice the data to obtain buildings in clusters of median size=250000 and storey height 15. But first, some more visualization graphs 

```{r echo=FALSE, message=FALSE}



ggplot(data = green_buildings) + 
  geom_point(mapping = aes(x = size, y = Rent), colour = 'green',alpha = 6/10) +
  geom_vline(data=green_buildings, aes(xintercept = median(green_buildings$size)), colour="red")


ggplot(data = non_green_buildings) + 
  geom_point(mapping = aes(x = size, y = Rent), colour = 'light blue') +
  geom_vline(data=non_green_buildings, aes(xintercept = median(non_green_buildings$size)), colour="red")


ggplot(data = green_buildings) + 
  geom_point(mapping = aes(x = stories, y = Rent), colour = 'green',alpha = 6/10) +
geom_vline(data=green_buildings, aes(xintercept = median(green_buildings$stories)), colour="red") 


ggplot(data = non_green_buildings) + 
  geom_point(mapping = aes(x = stories, y = Rent), colour = 'light blue') +
geom_vline(data=non_green_buildings, aes(xintercept = median(non_green_buildings$stories)), colour="red") 



```



```{r echo=FALSE, message=FALSE}

size_sum = Building_Data %>%
  group_by(cluster)  %>%  # group the data points by model name
  summarize(size2 = median(size))  



ggplot(size_sum, aes(x=reorder(cluster, size2), y=size2)) + 
  geom_bar(stat='identity') + 
  coord_flip()+labs(x = "Cluster",y='Median Size')

size_sum = Building_Data %>%
  group_by(cluster)  %>%  # group the data points by model name
  summarize(size2 = median(stories))  



ggplot(size_sum, aes(x=reorder(cluster, size2), y=size2)) + 
  geom_bar(stat='identity') + 
  coord_flip()+labs(x = "Cluster",y='Median Stories')



```

### We see that not all clusters have a median size of 250,000 and stories 15. We should consider that as a factor and shrink our data  


### Finding average median among green buildings of stories and size  

```{r echo=FALSE, message=FALSE}


average_median_stories=(aggregate(green_buildings$stories~green_buildings$cluster, data=green_buildings, median))


average_median=(aggregate(green_buildings$size~green_buildings$cluster, data=green_buildings, median))


```
```{r}


BD3=green_buildings[(average_median[2]>150000) & (average_median[2]<350000) & (average_median_stories[2]>11) & (average_median_stories[2]<30), ] 



```
```{r echo=FALSE, message=FALSE}


ggplot(data=BD3, aes(BD3$size)) + 
  geom_histogram() +
geom_vline(data=BD3, aes(xintercept = median(BD3$size)), colour="red") +
labs(x = "Cluster",y='Median Size of Green buildings')


ggplot(data=BD3, aes(BD3$stories)) + 
  geom_histogram() +
geom_vline(data=BD3, aes(xintercept = median(BD3$stories)), colour="red") +
labs(x = "Cluster",y='Median Stories of Green buildings')



#We see that the size is not equal among the data and the median is in 128838sq feet, Hence we cannot take this data to perform our analysis. We need to take 


```
```{r echo=FALSE, message=FALSE}

ggplot(data = BD3) + 
  geom_point(mapping = aes(x = size, y = Rent, color = green_rating),colour='green') +
  geom_vline(data=BD3, aes(xintercept = median(BD3$size)), colour="red") 

ggplot(data = BD3) + 
  geom_point(mapping = aes(x = stories, y = Rent, color = green_rating),colour='green') +
  geom_vline(data=BD3, aes(xintercept = median(BD3$stories)), colour="red") 


```


```{r}

ggplot(data=BD3, aes(BD3$Rent)) + 
  geom_histogram() +
geom_vline(data=BD3, aes(xintercept = median(BD3$Rent)), colour="red") 

median(BD3$Rent)


```


#### From this we see that the rent is 31.45  


### Now let us calculate for non-green buildings in the same clusters  


```{r}

size_sum = BD3 %>%
  group_by(cluster)  %>%  # group the data points by model name
  summarize(size2 = median(size))  



ggplot(size_sum, aes(x=reorder(cluster, size2), y=size2)) + 
  geom_bar(stat='identity') + 
  coord_flip()


```



### We see that the stories is 15, size is 250,00 hence this is appropriate data to figure out the RENT  

### The median rent is 31.45




```{r echo=FALSE, message=FALSE}



BD2_non_green=Building_Data[which(!Building_Data$green_rating == 1), ] 


BD3_non_green=BD2_non_green[(average_median[2]>150000) & (average_median[2]<350000) & (average_median_stories[2]>11) & (average_median_stories[2]<30), ] 



```
```{r echo=FALSE, message=FALSE}

ggplot(data=BD3_non_green, aes(BD3_non_green$Rent)) + 
  geom_histogram() +
geom_vline(data=BD3_non_green, aes(xintercept = median(BD3_non_green$Rent)), colour="red") 

median(BD3_non_green$Rent)

```


## Calculating :

### Median Rent for green buildings in clusters with median size = 250000 and median stories 15: 31.45  


### Median Rent for non-green buildings in clusters with median size = 250000 and median stories 15: 25.48

### Hence difference is : 31.45-25.48 = 5.97;  5.97*250,000=1,492,500  

### Amount taken to construct 5,000,000;  Hence it is : 5,000,000/1,492,500  
### **3.35 years to gain the money back**


### But let us see the median leasing rate : #92.285  


```{r echo=FALSE, message=FALSE}


ggplot(data=BD3, aes(BD3$leasing_rate)) + 
  geom_histogram() +
geom_vline(data=BD3, aes(xintercept = median(BD3$leasing_rate)), colour="red") 


```


### 3.35/92.285  
### = 3.63 years


